\documentclass[letterpaper,12pt]{article}
\setlength{\headheight}{15pt}
\setlength{\marginparwidth}{0pt}
\setlength{\marginparsep}{0pt} % width of space between body text and margin notes
\setlength{\evensidemargin}{0.125in} % Adds 1/8 in. to binding side of all 
% even-numbered pages when the "twoside" printing option is selected
\setlength{\oddsidemargin}{0.125in} % Adds 1/8 in. to the left of all pages when "oneside" printing is selected, and to the left of all odd-numbered pages when "twoside" printing is selected
\setlength{\textwidth}{6.375in} % assuming US letter paper (8.5 in. x 11 in.) and side margins as above
\raggedbottom
\setlength{\parskip}{\medskipamount}


\usepackage{amsmath, amsthm, amssymb, fancyhdr, enumitem, tikz, pgfplots, float}

\pgfplotsset{compat=1.18}

\pagestyle{fancy}
\lhead{Applicant Update --- Jonathan Ma}
\begin{document}
\paragraph{}Hey, thanks for reviewing my application. I've got some updates I'd like to share.

\paragraph{}Since submitting my application, I've been fortunate enough 
to have been extended a Summer Internship (2023) offer from \textbf{Chatham Financial}, 
a financial firm in Kennett Square, Pennsylvania. The position is in Software Engineering / Development.


\paragraph{}I've also begun a (paid) position at the \textbf{UD Artificial Intelligence Center of Excellence}, where
I'm working with Dr. Pinar Kullu, and Dr. Yao Yu to create an application which can demo
their research for peer review. Dr. Yu's project involves analyzing runoff risk in the Great Lakes,
and it's implemented using Deep Neural Networks. 

\paragraph{}My job is to essentially create a website to demo Dr. Yu's project, and in the future,
we'd like to provide Dr. Yu's project as a service to the US Government, and agricultural workers across
the world. 

\paragraph{}Again, thank you for reviewing my application, and thanks for considering these updates.


\paragraph{}Cheers,



Jonathan

\paragraph{} \textbf{P.S.} I've also uploaded a copy of my resume to the Princeton Applicant Portal.
\end{document}



