\documentclass[letterpaper,12pt]{article}
\setlength{\headheight}{15pt}
\setlength{\marginparwidth}{0pt}
\setlength{\marginparsep}{0pt} % width of space between body text and margin notes
\setlength{\evensidemargin}{0.125in} % Adds 1/8 in. to binding side of all 
% even-numbered pages when the "twoside" printing option is selected
\setlength{\oddsidemargin}{0.125in} % Adds 1/8 in. to the left of all pages when "oneside" printing is selected, and to the left of all odd-numbered pages when "twoside" printing is selected
\setlength{\textwidth}{6.375in} % assuming US letter paper (8.5 in. x 11 in.) and side margins as above
\raggedbottom
\setlength{\parskip}{\medskipamount}


\usepackage{amsmath, amsthm, amssymb, fancyhdr, enumitem, tikz, pgfplots}

\pgfplotsset{compat=1.18}

\pagestyle{fancy}
\lhead{CISC220H - Programming 6 Report}
\begin{document}
We use two heaps to satisfy our specification in $O( \mathrm{log}(n))$ time.
Our first heap keeps track of our lower-half-valued elements, and the second heap keeps track of the
greater-half-valued elements. We maintain an invariant that these two heaps are at most one element
apart from each other in size, thus easily allowing us to find the median of our inserted elements.



We maintain a running median, which we set to 0 to begin.


We insert elements one by one, and at each inserted element, we check several cases as follows:


\begin{itemize}
    \item If our second heap is larger than our first heap:
        \begin{itemize}
            \item If the inserted element is greater than our current median, we move the top element
                from the first heap to the second heap, and push the inserted element into the second heap.
            \item Otherwise, we push the inserted element into the first heap.
            \item We let the median be equal to the top element of the first heap.
        \end{itemize}
    \item If our first heap is larger than our second heap:
        \begin{itemize}
            \item If the inserted element is less than the current median, we move the top element
                from the first heap to the second heap, and push the inserted element into the first heap.
            \item Otherwise, we push the inserted element into the second heap.
            \item We let the median be equal to the top element of the first heap.
        \end{itemize}
    \item If both heaps are equal:
        \begin{itemize}
            \item If the inserted element is less than the current median, we push it into the first heap,
                and let the median be equal ot the top of the first heap.
            \item Otherwise, we push the inserted element into the second heap, and let
                the median be equal to the top element of the second heap.
        \end{itemize}
\end{itemize}
\end{document}



